---
title: "R Notebook 2025: NWOpen-API"
output: 
  html_document:
    toc: true
    toc_depth: 5
    toc_float: TRUE

---

### Abstract

This r notebook can retrieve the grant information from NWO (Nederlandse organisatie voor Wetenschappelijk Onderzoek) between 2 specific dates from an organisation (in this example ‘Universiteit Leiden’) using their API. The API returns results in JSON format. The data schema is available via [this PDF file](https://www.nwo.nl/en/how-to-use-the-nwopen-api).

The r notebook output will be multiple Excel files in a sub-folder of DATA_GRANTS based on your query.




# Introduction

De Nederlandse organisatie voor Wetenschappelijk Onderzoek (NWO) projectendatabank raadplegen via de NWOpen API om de projecten waar LEI aan is verbonden uit de selectie die beschikbaar is vanuit de [projectendatabank](https://www.nwo.nl/projecten). De projectendatabank interface gaat terug tot 1993 terug, maar lijkt in het begin onvolledig. Informatie over de data selectie via de API staat onder data.

ZonMW heeft een separate projectendatabank via: https://projecten.zonmw.nl/nl

Voor meer informatie over de API van NWO zie het PDF-bestand "NWOpen-API" via: https://www.nwo.nl/gebruik-nwopen-api


# Setup

## Load packages

```{r, PACKAGES, echo = TRUE, eval = TRUE}

# Packages needed
  package.list <- c("conflicted", "reprex", "tidyverse", "rmarkdown",
                    "httr2", "data.table", "jsonlite", "vctrs", "openxlsx")


# Install missing packages 
  new.packages <- package.list[!(package.list %in% installed.packages()[,"Package"])]
  if(length(new.packages)) install.packages(new.packages)

  
# Load packages needed (and subdue)
  invisible(lapply(package.list, library, character.only = TRUE))


```



# Data

De data is afkomstig uit ISAAC, het administratiesysteem van NWO. De informatie wordt ook gepubliceerd via NWO’s projectendatabank op de website: https://www.nwo.nl/projecten . De data heeft de volgende beperkingen:

* API data is incompleet: 
  + De API levert informatie die wordt geleverd bevat lopende projecten en projecten die maximaal 5 jaar geleden zijn afgesloten. (niet de volledige projectendatabank die teruggaat tot ± 1993; beginjaren zijn onvolledig).
  + projecten die door de projectleider met “niet publiceren” zijn aangemerkt, of projecten waarvan de call door NWO als “niet publiceren” is aangemerkt worden niet door de API zullen worden teruggegeven.
* project_id ≠ Grant_id (2024)
  + Vanaf augustus 2024 krijgen NWO projecten een nieuwe Grant ID is een persistent identifier die linkt naar de projectpagina op de website van NWO met informatie over de subsidie met de vaste prefix: https: //doi.org/10.61686/… .
  + NWO vraagt de projectleider de grant ID te vermelden bij alle publicaties of preprints, data en software.
* Op dit moment lopen er diverse projecten binnen NWO om persistent identifiers toe te voegen. Er wordt gewerkt aan het toevoegen van DOI, ORCID en ROR. De NWOpen-API is hier op voorbereid en zal de informatie teruggeven wanneer de systemen van NWO met de betreffende identifiers zijn verrijkt.
* beperking van 100 projecten per query
  + 1993-01-01 t/m 2024-12-31 data begint in 2013.


## Get data from NWOpen-API


### API base URL

```{r, API URL, echo = TRUE, eval = TRUE}

# URL NWOpen-API
  base_url <-"https://nwopen-api.nwo.nl"

```



### API query: Iterative

```{r, API QUERY: ITERATIVE & PROCESSING RESPONSE, echo = TRUE, eval = TRUE}

### API request and iterative response processing to a data frame (NWO_data_projectendatabank)

# Dates is given as ISO 8601 extended format YYYY-MM-DD
  NWO_responses <- 
    request(base_url) |>
    req_url_path_append ("NWOpen-API/api/Projects") |>
    req_url_query(
      rs_start_date = "2020-01-01",
      re_start_date = "2024-12-31",
      organisation = "\"Universiteit Leiden\"",
#      per_page = 10     # reduces repsone length for testing (in combination with max_reqs)
      ) |> 
  
    req_perform_iterative(
      next_req = iterate_with_offset(
        "page", 
        resp_pages = function(resp) resp_body_json(resp)$meta$pages
        ),
      max_reqs = Inf # use for testing max_reqs = 3 if you want to limit your max requests
        ) |>
    resps_successes() |>
    resps_data(\(resp) resp_body_json(resp)$projects)



# Convert NWO_responses to JSON (JavaScript Object Notation)
  NWO_responses_JSON <- toJSON(NWO_responses)


# Convert NWO_responses from JSON to data frame
  NWO_data_projectendatabank <- fromJSON(NWO_responses_JSON)

```



# Processing API response


## Create directory

```{r, BEFORE PROCESSING CREATE FOLDERS, echo = TRUE, eval = TRUE}

# Create Folders: creates a DATA_GRANTS directory (if not already present) with a sub-folder
  ifelse(!dir.exists("DATA_GRANTS"), dir.create("DATA_GRANTS"), "Folder exists already")
  ifelse(!dir.exists(file.path("DATA_GRANTS", "NWO_GRANTS")), dir.create(file.path("DATA_GRANTS", "NWO_GRANTS")), "Folder exists already")

```



## Processing the response
<a id="members"></a>

### PROJECTS: members

```{r, PROCESSING API RESPONSE: MEMBERS, echo = TRUE, eval = TRUE}

### Processing api response of project members (Personenbestanden) with project_id #
  
# Selects project_members and project_id
  NWO_projects_members_all <- 
    select(NWO_data_projectendatabank, 
           project_id,
           project_members,
           )


# Unnest project_members from NWO_projects_members_all
  NWO_data_personen <- NWO_projects_members_all %>% unnest(project_members)


# Filters NWO_projects_members_all by role
  NWO_members_main.applicant <- subset(NWO_data_personen, role == "Main Applicant") # ≠ # projects
  NWO_members_Project.leader <- subset(NWO_data_personen, role == "Project leader") # still lose 1 project
  NWO_members_Researcher <- subset(NWO_data_personen, role == "Researcher") # solves missing projects


# Combining previous filters (order matters to get the role = "Main Applicant" for most projects)
  NWO_members_sort <- do.call("rbind", list(NWO_members_main.applicant, NWO_members_Project.leader, NWO_members_Researcher))  


# Deduplicate the combined file          
  NWO_data_personen_hoofdaanvrager <- NWO_members_sort[!duplicated(NWO_members_sort$project_id),]



# Save: create excel files with the project member (with project_id # column)
  write.xlsx(NWO_data_personen_hoofdaanvrager,  "DATA_GRANTS/NWO_GRANTS/NWO_LEI_projecten_personen_hoofdaanvragers.xlsx") # with only "Main Applicant"
  write.xlsx(NWO_data_personen, "DATA_GRANTS/NWO_GRANTS/NWO_LEI_projecten_personen.xlsx") # with all project_members

```
<a id="projects"></a>


### PROJECTS: projects (grants)

```{r, PROCESSING API RESPONSE: PROJECTS, echo = TRUE, eval = TRUE}

### Processing api response of projects (Projectenbestand)

# Selects projects information
  NWO_projecten <- select(NWO_data_projectendatabank,
                          project_id, # ≠ grant id
                          title,
                          funding_scheme_id,
                          funding_scheme,
                          department,
                          sub_department,
                          start_date,
                          end_date,
                          summary_nl,
                          summary_en,
                          grant_id, # DOI as of ± September 2024 (if the test fails dates are outside the implementation range) ≠ project_id
                          )


# Combining "NWO_projecten" and "NWO_data_personen_hoofdaanvrager" by project_id
  NWO_data_projecten <- left_join(NWO_projecten, NWO_data_personen_hoofdaanvrager, by = "project_id")



# Save: create excel files with projects (grants)
  write.xlsx(NWO_projecten, "DATA_GRANTS/NWO_GRANTS/NWO_LEI_projecten.xlsx")
  write.xlsx(NWO_data_projecten, "DATA_GRANTS/NWO_GRANTS/NWO_LEI_projecten_en_hoofdaanvrager.xlsx") # with "Main Applicant"

```
<a id="products"></a>


### PROJECTS: products

```{r, PROCESSING API RESPONSE: PRODUCTS, echo = TRUE, eval = TRUE}

### Processing api response of products (onderzoeksoutput) and project_id

# Select products and project_id
  NWO_projects_products <- 
    select(NWO_data_projectendatabank, 
           project_id,
           products,
           )


# Unnest products from NWO_projects_products and authors
  NWO_output <- NWO_projects_products %>% unnest(products)
  NWO_data_onderzoeksoutput <- NWO_output %>% unnest(authors)
  

# Select Product information without author information
  NWO_data_output <- select(NWO_output,
                          project_id,
                          type,
                          title,
                          sub_title,
                          year,
                          url_open_access, # = doi as url
                          journal_title,
                          isbn,
                          )


# Select Product information with author information  
  NWO_data_onderzoeksoutput <- select(NWO_data_onderzoeksoutput,
                          project_id,
                          type,
                          title,
                          sub_title,
                          year,
                          url_open_access, # = doi as url
                          journal_title,
                          isbn,
                          last_name,
                          initials,
                          first_name,
                          prefix,
                          dai, # (depending on the scope of the query the data will not contain this column)
                          )


# List as character (see Notes below)
#  NWO_data_onderzoeksoutput  <- NWO_data_onderzoeksoutput |> mutate(across(where(is.list),as.character))
#  NWO_data_output  <- NWO_data_output |> mutate(across(where(is.list),as.character))

  

# Save: create excel files with products (research output) with project_id # column
  write.xlsx(NWO_data_output, "DATA_GRANTS/NWO_GRANTS/NWO_LEI_onderzoeksoutput.xlsx")  
  write.xlsx(NWO_data_onderzoeksoutput, "DATA_GRANTS/NWO_GRANTS/NWO_LEI_onderzoeksoutput_met_auteurs.xlsx")  

```



# Notes

The data for the response to your query is stored in NWO_data_projectendatabank. You do not need to run all code from "Processing the different elements" onwards. Points of failure during the processing of the response could be the following:

* [PROJECTS: projects (grants)](#projects)
  + If the end date of your query is ± before September 2024 it doesn't contain the grant_id and this element will fail.
* [PROJECTS: products](#products)
  + If your query is small and near the actual time you are running this query, there will be no research output and this chunk will fail.
  + Depending on the scope of the query the data will have no authors with a DAI (Digital Author Identifier)
  + When exporting data frame to openxlsx: Error in x[is.na(x)] <- na.string : replacement has length zero occurs use the provides "List as character" solution to still export the data.frame (occurs when no dates are used).


Also note:

* [PROJECTS: members](#members)
  + The data in the organisation  column is not always clear (the data is identical to the data the main grant applicant wrote on the application form).



